### 2023-01-18 paper/main.tex

\documentclass[10pt,sigconf,letterpaper,anonymous,nonacm]{acmart}

\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{tikzscale}
\usetikzlibrary{positioning}
\usetikzlibrary{shapes.multipart}
\usetikzlibrary{graphs}
\usetikzlibrary{trees}
\usetikzlibrary{quotes}

\graphicspath{ {../viz/} }

\title{Tempus: Probabilistic Network Latency Verifier}
\author{WIP}

\begin{document}
\begin{abstract}
    To combat problems and bugs that a given network might have, network verifiers have emerged as 
    one of the promising solutions. 
    State-of-the-art network verifiers mainly focused on evaluating qualitative properties under 
    various scenarios of network failures, such as reachability under k-link failure. 
    However, as modern networks evolved and performance need becomes more stringent -- often 
    expressed in terms of Service Level Agreement (SLA) -- there is a need to evolve network 
    verifiers to also reason about quantitative performance properties. 
    Works in quantitative network verifiers that has arose in recent years mainly focused on one 
    side of the network performance metric: bandwidth and load violation properties. 
    Questions about the other side of network performance metric, latency, were left unanswered.

    In this work, we introduce a verifier framework, Tempus, that can answer the probability of 
    a given temporal property being true given latency distributions of individual links and 
    nodes in the network. 
    Early evaluation shows that Tempus can verify bounded reachability property -- the probability
    that the average delay of packets traversing from a source to destination node is below $T$
    time unit -- by only adding a fraction of the state exploration overhead introduced by the 
    qualitative verifier it's built upon.
\end{abstract}

\maketitle

\input{1_intro.tex}

\input{2_overview.tex}

\input{3_functional.tex}

\input{4_temporal.tex}

\input{5_eval.tex}

\input{6_related.tex}

\input{7_future.tex}

\section{Conclusion}
\begin{itemize}
    \item Tempus is a scalable latency SLA verifier. It could probabilistically verify in the order of minutes
    \item Consolidation and memoization is an effective optimization strategy, although the exact magnitude of 
        its effectiveness will depend on the network topology, specifically on the amount of paths and their 
        length.
\end{itemize}

\bibliography{cit}
\bibliographystyle{ieeetr}
\end{document}