### 2023-02-08 paper/main.tex

\documentclass[10pt,sigconf,letterpaper,anonymous,nonacm]{acmart}

\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{tikzscale}
\usetikzlibrary{positioning}
\usetikzlibrary{shapes.multipart}
\usetikzlibrary{graphs}
\usetikzlibrary{trees}
\usetikzlibrary{quotes}

\graphicspath{ {../viz/} }

\newcommand{\tool}[0]{Tempus}

\title{Probabilistic Network Latency Verification}
\author{WIP}

\begin{document}
\begin{abstract}
    To combat configuration errors that a given network might have, network verifiers have emerged as 
    one of the promising solution. 
    State-of-the-art network verifiers mainly focused on evaluating qualitative properties under 
    various scenarios of network failures, such as reachability and loop detection.

    However, as modern networks evolved and performance need becomes more stringent -- often 
    expressed in terms of Service Level Agreement (SLA) -- there is a need to evolve network 
    verifiers to also reason about quantitative performance properties. 
    Works in quantitative network verifiers that has arose in recent years mainly focused on one 
    side of the network performance metric: bandwidth and load violation properties. 
    Questions about the other side of network performance metric, latency, were left unanswered.

    In this work, we introduce a verification framework, \tool, that can answer the probability of 
    a given temporal property being true given latency distributions of individual links and 
    nodes in the network. 
    We use the convolution operation on these probability distributions to calculate the end-to-end 
    path latency distribution, and answer property by analyzing the statistical property of this 
    resulting distribution.
    We also introduce two effective optimization techniques to exploit the symmetry in the verification 
    process to reduce the amount of convolution operation needed.

    Our evaluation shows that \tool can verify temporal property 4 orders of magnitude faster on average than 
    our baseline, which is devoid of all optimization.
    We also identify that highly symmetrical topology, like Fat Tree, benefits from our optimization the most 
    and can cut the amount of convolution up to 6 orders of magnitude.
\end{abstract}

\maketitle

\input{1_intro.tex}

\input{2_overview.tex}

\input{3_functional.tex}

\input{4_temporal.tex}

\input{5_opt.tex}

\input{6_eval.tex}

\input{8_future.tex}

\input{7_related.tex}

\section{Conclusion} \label{sec:conc}
\begin{itemize}
    \item Tempus is a scalable latency SLA verifier. It could probabilistically verify in the order of minutes
    \item Consolidation and memoization is an effective optimization strategy, although the exact magnitude of 
        its effectiveness will depend on the network topology, specifically on the amount of paths and their 
        length.
\end{itemize}

\bibliography{cit}
\bibliographystyle{ieeetr}
\end{document}