### 2023-02-13 paper/main.tex

\documentclass[10pt,sigconf,letterpaper,anonymous,nonacm]{acmart}

\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{tikzscale}
\usetikzlibrary{positioning}
\usetikzlibrary{shapes.multipart}
\usetikzlibrary{graphs}
\usetikzlibrary{trees}
\usetikzlibrary{quotes}

\graphicspath{ {../viz/} }

\newcommand{\tool}[0]{Tempus}

\title{Probabilistic Network Latency Verification}
\author{WIP}

\begin{document}
\begin{abstract}
    As modern networks evolved and performance need became more stringent, often 
    expressed in terms of Service Level Agreement (SLA), there is a rising need to verify that these 
    agreements hold true.
    Designing a network verifier that can practically reason about quantitative properties proves to be 
    a unique challenge, since it needs to be able to expressively model many quantitative phenomenon, like 
    congestion control and load balancing scheme, while being reasonably performant and not 
    requiring too many inputs from its user. 
    In this work, we introduce a network latency verification framework, \tool, that could probablistically 
    verify temporal properties of a network given latency measurements of its components. 
    We also introduce two effective optimization techniques to exploit the symmetry in the verification 
    process to significantly reduce the verification runtime.
    Our evaluation shows that \tool\ can verify temporal property 90\% faster on average than 
    our baseline, which is devoid of all optimization.
    We also identify that highly symmetrical topology, like Fat Tree, benefits from our optimization the most 
    and can cut the amount of convolution up to 6 orders of magnitude.
\end{abstract}

\maketitle

\input{1_intro.tex}

\input{2_overview.tex}

\input{3_functional.tex}

\input{4_temporal.tex}

\input{5_opt.tex}

\input{6_eval.tex}

\input{8_future.tex}

\input{7_related.tex}

\section{Conclusion} \label{sec:conc}
\begin{itemize}
    \item Tempus is a scalable latency SLA verifier. It could probabilistically verify in the order of minutes
    \item Consolidation and memoization is an effective optimization strategy, although the exact magnitude of 
        its effectiveness will depend on the network topology, specifically on the amount of paths and their 
        length.
\end{itemize}

\bibliography{cit}
\bibliographystyle{ieeetr}
\end{document}